The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2011-11-29 smurf_paper.tex
Gap filling, apodisation and padding

DSB

Discuss gap filling, ``new'' window function method, also apodization
and start/end polynomial continuity stuff. Should reference CMB
literature -- in particular look at \citet{stompor2002}.

>>>>

The evaluation of the FLT model (Section~\ref{sec:flt}) involves taking
the Fast Fourier Transform of each bolometer time series. Data that has
been flagged as bad for any reason (for instance, due to the presence of
spikes, steps, or unusual common-mode signal) needs to be excluded from
this FFT. Two method are available:

\begin{itemize}
\item Gap-filling: each contiguous block of bad data samples is filled
with artificial data before taking the FFT. A least squares linear fit is
performed to the 50 samples preceding the block, and a similar fit is
performed to the 50 samples following the block. These are used to
estimate the expected values at the start and end of the block of bad
values. The bad values are then replaced by linear interpolation between
the expected start and end values. Gaussian noise is added with a
standard deviation equal to the mean of the RMS residuals in the two
fits.
\item Masking: each bad data value is replaced by zero, and an associated
mask is created that holds 1.0 for each good sample in the time stream and
0.0 for each bad sample. The masked data and the mask are then each filtered
in the same way using FFTs, and the filtered data is finally normalised by
dividing it by the filtered mask.
\end{itemize}

In addition to replacing bad samples before the FFT, it is also necessary
to ensure that the data values at the start and end of the time series
are similar. Since a Fast Fourier transform treats the data as a single
cycle in an infinitely repeating waveform, any large difference between
starting and ending value will effectively introduce sudden steps at the
start and end of each cycle, causing unwanted oscillations (ringing) in
the transform. Another consequence of the cyclic nature of the FFT is
that features at one end of the time series can affect the filtered
values at the other end of the time series. Two methods are available to
avoid these two problems:

\begin{itemize}
\item Apodisation: a number of samples at the start and end of each
bolometer time series are multiplied by a cosine function in order to
roll the data values off smoothly to zero. The default number of samples
modified at each end of the time series is given by half the ratio of the
sampling frequency to the lowest retained frequency. In addition, each
end of the time stream is padded with double this number of zeros. This
method is illustrated in Fig.~\ref{fig:pad2}.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pad2.pdf}
\caption{The black curve shows a bolometer time stream, padded
with zeros. The red curve shows the time stream after apodisation.}
\label{fig:pad2}
\end{figure}

\item Padding with artificial data: instead of padding with zeros, each
time stream is padded with artificial data that connects the two ends of
the data stream smoothly and includes Gaussian noise. No apodisation is
performed. The number of samples of padding at each end is again equal to
the ratio of the sampling frequency to the lowest retained frequency.
This is illustrated in Fig.~\ref{fig:pad1}.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pad1.pdf}
\caption{The black curve shows the same bolometer time stream as in
Fig.~\ref{fig:pad2}, againpadded with zeros. The red curve shows the time stream
after padding with artificial data.}
\label{fig:pad1}
\end{figure}

\emph{Should reference CMB literature -- in particular look at
\citet{stompor2002}}.
